General constraints on magic distillation protocols from statistical mechanics

### 2021-06-06 magic_paper.tex
Merge choices made.

We define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(W_\rho)$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(W_\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_vs_eps_a.pdf}
    \caption{\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]} Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $H_\alpha$ is increasing in $\alpha$ and $\epsilon$ \nick{should be easy to show analytically}. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. Specifically, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ has a unique solution, which is the \nick{probably irrational} solution of $3^\alpha - 2^{3-\alpha} = 1$.
    }
    \label{fig:lcs}
\end{figure}

@@

We now show the following simple equivalence.
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.
\begin{proof} If we have $H_\alpha (W_\rho(\x)) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n=a_n/b_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}
Therefore Wigner negativity coincides with negativity of a Renyi entropy.

>>>>

We define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(W_\rho)$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(W_\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
Inducing any magic state $\rho$ with depolarising noise, $\rho(\epsilon) = (1-\epsilon)\rho + \epsilon \sigma$ results in $H_{\alpha}\left(\rho(\epsilon)\right) \leq H_\alpha(\rho)$ because there exists a free operation $\E \in \O$ that transforms $\rho(\epsilon) = \E(\rho)$ and $H_\alpha$ is Schur-concave.

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_vs_eps_a.pdf}
    \caption{\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]} Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $H_\alpha$ is increasing in $\alpha$ and $\epsilon$. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. Specifically, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ has a unique solution, which is the \nick{probably irrational} solution of $3^\alpha - 2^{3-\alpha} = 1$.
    }
    \label{fig:H}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/D_vs_eps_a.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $D_\alpha$ is decreasing in $\alpha$ and $\epsilon$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:D}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $H_\alpha$ is decreasing in $\epsilon$ (not Schur-conocave for $\alpha < 1$) and increasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Htest}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/D_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $D_\alpha$ is increasing in $\epsilon$ (not Schur-convex for $\alpha < 1$) and decreasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Dtest}
\end{figure}

\begin{lemma}
	Let $\alpha \geq \beta > 1$ where $\alpha, \beta$ are even$/$odd.
	Then,
	\begin{equation}
		H_\alpha(\x) \geq H_\beta(\x) \text{ and } D_\alpha(\x || \r) \leq D_\beta(\x || \r)
	\end{equation}
	for any quasi-distribution $\x \in \mathbb{R}^N$ and probability distribution $\r \in \mathbb{R}^N$.
\end{lemma}
\begin{proof}
	For $\alpha \geq \beta > 1$, we have
	\begin{equation}
		0 > \frac{1}{1-\alpha} \geq \frac{1}{1-\beta},
	\end{equation}
	and $w_i^\alpha = |w_i|^\alpha \leq |w_i|^\beta = w_i^\beta$.

	\nick{how to proceed?}

	For any $\r$ with positive rational components, we have $D_\alpha(\x || \r) \leq D_\beta(\x || \r)$ due to~\cref{thm:H2D}.
\end{proof}

We can prove additivity of $H_{\alpha}$ on quantum states.

\begin{lemma}\label{H_add}
	For any quantum states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho'),
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}

@@

We now show the following simple equivalence.
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.

\begin{proof} If we have $H_\alpha (W_\rho(\x)) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n=a_n/b_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
\nick{In this step, we also need to show that $H_\alpha$ is increasing in $\alpha$}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}

Therefore Wigner negativity coincides with negativity of a Renyi entropy.
